---
title: "Spatial Econometrics"
author: "Grzegorz Krochmal and Ewelina Osowska"
date: ""
output:
  html_document:
    toc: true
---
<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 120)
```

### Introduction

The following analysis ... 

### Dataset 
<!-- opis zmiennych, jakieś statystyki może podstawowe, jakie wypełnienie missingów -->
https://stackoverflow.com/questions/27928/calculate-distance-between-two-latitude-longitude-points-haversine-formula/23095329#23095329

The dataset used in the analysis is created based on ...

```{r}
df_final <- readxl::read_excel("df_final.xlsx")
```

### Initial analysis
<!-- mapka ze średnimi cenami vs dzielnica, średni czynsz vs dzielnica - z pythona, Grzesiek chciał zrobić -->



### Empirical analysis

In order to estimate housing prices, we run several models and then compare them with respect to prediction power. There are two types of models which we consider: regular OLS regression and spatial models. 

#### OLS regression


#### Modele przestrzenne
<!--macierz wag, test I Morana, jakość modeli: AIC, BIC, LogLik-->

#### Non-parametric model



